Type-Safe Modular Parsing

### 2017-05-29 SLE_17/sections/Inheritance.tex
A bit.

@@

@@

To illustrate how to modularly add new language constructs and the
corresponding parsing code, we continue with the expression language
of literals and additions in Section~\ref{subsec:packratparsing}. We
introduce variables as a new case:

\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= ...
   \alt <ident>
\end{grammar}

It is easy to extend the corresponding OO AST in a modular way:

@@

Since we already have the parser for literals and additions, we would
like to build the new parser by reusing the old one. In Scala we can take advantage of inheritance for such reuse. Specifically,
the new parser can be defined in an enclosing trait that extends
the old one. Here one may quickly come up
with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
with \inlinecode{pExpr}:

@@

\subsection{Multiple Extensions and Independent Extensibility}
A nice feature of Scala is its support for a form of multiple
inheritance.  Scala has a linearized-style mixin inheritance for
traits~\cite{odersky2004overview}. This can be very helpful when
composing several languages, and to achieve independent
extensibility~\cite{odersky2005independently}.
Suppose now we want to compose the parsers for expressions
from pre-defined languages \inlinecode{LanguageA} and
\inlinecode{LanguageB} using alternative. The new parser can be built
by inheriting both parsers at the same time:

\lstinputlisting[linerange=46-50]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}. Under multiple inheritance, it can be used to distinguish the methods of the same name. Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.

@@

\paragraph{Conflicts and/or ambiguity} As we demonstrated before, the longest match alternative combinator is useful when composing parsers. However, ambiguous grammars still cause subtle issues in modular parsing. For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} really depends on the implementation of the alternative combinator, because the results can not be prioritized by length. Since the composition is handled explicitly by a user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.

>>>>

\begin{comment}

@@

\end{comment}

@@

Part of the modular parsing problem is how to obtain an extensible parser. It
is natural to make use of OO ASTs because adding new data constructs is cheap for
them. Hence we have used OO traits and inheritance to represent the AST in the last
section. Furthermore, we would like to write extensible parsing code on extensions of
a grammar. That is to say, new extensions would not require modification on the existing code,
and we can even reuse the old code.

To illustrate such extensibility, we continue with the old
example, and introduce variables as a new case.
It is easy to extend the corresponding OO AST in a modular way:

@@

For code reuse, we want to define a new parser that extends
the old one. Here one may quickly come up
with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
with \inlinecode{pExpr}:

@@

\vspace{-.05in}
\paragraph{Independent Extensibility}
A nice feature of Scala is its support for the linearized-style multiple
inheritance on traits~\cite{odersky2004overview}. This can be very helpful when
composing several languages, and to achieve independent
extensibility~\cite{odersky2005independently}. Later we introduce language components
in Section~\ref{subsec:language-component} as a nice pattern.

In the above code, we use the \lstinline{super} keyword to refer to the old \inlinecode{pExpr}.
With multiple inheritance, Scala has a special syntax
\lstinline{super[T].x}, called \textit{static super reference}. It refers to the type or
method \inlinecode{x} in the specified parent trait \inlinecode{T}. With this we can reuse
the old parsers even if they have the same name.

@@

\vspace{-.05in}
\paragraph{Conflicts and/or ambiguity} In a modular setting, conflicts and
ambiguity could be introduced to the grammar. In that case, the help parser combinators
can offer is quite restricted. Yet users can override those problematic methods to resolve
such conflicts, and rely on dynamic dispatch. We will discuss it in Section~\ref{subsec:overriding-rules}.
